Some examples in extended logic counting classes

### 2015-08-04 tex/case3.tex

Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$. In order to prove that $f\dotminus 1\in\E1$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
\begin{eqnarray*}
\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
\end{eqnarray*}
and
\begin{eqnarray*}
\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
\end{eqnarray*}
where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables. Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2. The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.

In this case, we mix both the strategies in cases 1 and 2. That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.

Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
\begin{multline*}
\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ \varphi_{m_i,<}((\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}),(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i}))\Big)\Big],
\end{multline*}
and let
\begin{multline*}
\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{p,<}(\w,\z)).
\end{multline*}
Let $\P$ and $\e$ be assignments to $\X$ and $\z$. Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
\begin{multline*}
\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
\end{multline*}
Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.

Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$. Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$. Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b). The proof of this is analogous to case 2. Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$

>>>>

Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$. In order to prove that $f\dotminus 1\in\E1$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
\begin{eqnarray*}
\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
\end{eqnarray*}
and
\begin{eqnarray*}
\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
\end{eqnarray*}
where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\lambda_i(j)} + c_{\lambda_i(j)} = a_{\lambda_i(j)}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables. Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2. 
\icristian{
Este orden total considera 	$\x,\z$ como un todo? Sería bueno recalcar esto.
}
The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.

In this case, we mix both the strategies in cases 1 and 2. That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
\icristian{
Esta intuición no se entiende nada. Mejorar al igual que las intuiciones anteriores.
}

Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
\begin{multline*}
\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ \varphi_{m_i,<}((\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}),(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i}))\Big)\Big],
\end{multline*}
and let
\begin{multline*}
\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{p,<}(\w,\z)).
\end{multline*}
Let $\P$ and $\e$ be assignments to $\X$ and $\z$. Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
\begin{multline*}
\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
\end{multline*}
Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.

Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$. Let $\P$ be the lexicographically smallest tuple of predicates 
\icristian{
Acá no esta bien definido que significa ``lexicographically smallest tuple of predicates''. También deberias hablar de ``$(\P, \z)$ be the lexicographically smallest tuple of predicates'' incluyendo a $\z$.
}
such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$. Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b). The proof of this is analogous to case 2. Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$

### 2015-08-04 tex/lemmafirst.tex

\begin{lemma} \label{first}
$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
\end{lemma}
\begin{proof}
Let $\A \in \Truc$. Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$. Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$. Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$. We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
\begin{enumerate}
\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$. That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$). Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$. Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$. We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$. Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$. But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent. That is, there's no structure $\A^\prime\in\Truc$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$. In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$. Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$. Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have. If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. On both cases, we have a contradiction.
\end{enumerate}
With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$

>>>>

\begin{lemma} \label{first}
$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
\end{lemma}
\begin{proof}
Let $\A \in \Truc$. Consider two cases: 
\icristian{
Acá yo explicaría cuales son los dos casos para entender el roadmap de la demostración.	
}
assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$. Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$. Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$. We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).

\begin{enumerate}
\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$. That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$). Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$. Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$. We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$. Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$. But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent. That is, there's no structure $\A^\prime\in\Truc$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$. In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
\icristian{Este argumento es uno de los principales. Si cambiamos el orden de la demostración (colocando la descomposicion DNF primero) esta demo se podría colocar en el Claim 5.1 y después reusar acá}

\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
\icristian{
Aqui yo recalcaría que 	$\A$ no modela la segunda parte de $\varphi^\prime(\x,\P)$.
}
\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$. Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_{\lambda(i)}$. Then $\A\models \neg P^\prime_{\lambda(i)}(\d_i)$, and so,
$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ 
from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have. If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P'_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. On both cases, we have a contradiction.
\icristian{
Esta demostración es mas clara si es directa, diciendo que para todo $\P^\prime \neq \P$ se cumple que $\A\models\exists\x\,\varphi(\x,\P^\prime)$ y $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
}
\end{enumerate}
With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$

### 2015-08-04 tex/th2.tex

Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
\begin{eqnarray*}
S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
\end{eqnarray*}
Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
$$
\varphi(x,X) = (X(x) \vee \neg X(x)).
$$
Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. As we supposed initially, let $h\in\E1$ such that $f_{all}-f_{\#3DNF} = h$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc co-NP}-complete. However, as we showed previously in Theorem \ref{decisionptime}, $L_h \in \textsc{P}$. Then, $\textsc{co-NP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
\end{proof}

>>>>

Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
\begin{eqnarray*}
S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
\end{eqnarray*}
Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
$$
\varphi(x,X) = (X(x) \vee \neg X(x)).
$$
Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. As we supposed initially, let $h\in\E1$ such that $f_{all}-f_{\#3DNF} = h$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc coNP}-complete. However, as we showed previously in Theorem \ref{decisionptime}, $L_h \in \textsc{P}$. Then, $\textsc{coNP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
\end{proof}

### 2015-08-04 tex/th4.tex

\begin{theo}
$\E1$ is closed under substraction by one.
\end{theo}
\begin{proof}
Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$. Without loss of generality, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
\begin{eqnarray*}
	f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$. Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily. First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
\begin{eqnarray*}
	\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
\end{eqnarray*}
This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
\begin{eqnarray*}
	\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
\end{eqnarray*}
If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$. Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$. Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$

>>>>

\begin{theo}
$\E1$ is closed under substraction by one.
\end{theo}
\begin{proof}
Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
\begin{eqnarray*}
	f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$. Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily. First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
\begin{eqnarray*}
	\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
\end{eqnarray*}
This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
\begin{eqnarray*}
	\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
\end{eqnarray*}
If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$. Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$. Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
